---
title: "Project 1"
author: "Maya Szymborski"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: right
editor: source
execute:
  error: true
  warning: false
  message: false
  cache: true
---

# Introduction

The goal of this project is to create reusable, generic functions for working with data from the U.S. Census Bureau's Public Use Microdata Sample (PUMS) API. The functions should be able to retrieve data from different survey years, variables, and geographic areas, and return the data consistently without manually repeating the same data-processing steps.  

The PUMS API provides psuedo person-level records, but the records do not represent exactly one person but can represent multiple people. The `PWGTP` variable is a person weight, which indicates how many people a particular record represents. Because of this, the functions need to account for these weights. 

I focused on the PUMS data from 2021 through 2024, and excluded Puerto Rico as stated in the project instructions. My approach was to first understand how the Census API returned data, then create helper functions to process the metadata and special variables, and finally build and test functions that could retrieve, combine, summarize, and plot the data. 

# Part 1: Data Processing

Part 1 of the project focuses on retrieving and processing the Census API's PUMS data.

```{r}
# attach packages 
library(httr)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(quantreg)
```

## Explore PUMS Census API

I began by making a request to the 2024 Census PUMS API using the example provided in the project instructions. This let me see what the returned data from the Census API looked like. 

Since the response was returned as a JSON, I converted it to an R object. The first row contained the variable names, so I saved that row as `column_names` before removing it from the data. I then converted the response to a tibble and assigned `column_names` to the tibble with `names()`. 

I then used those steps to create a helper function, `query_api`. This function takes an inputted URL, sends the request, converts the JSON response into an R object, uses the first row as the column names, and returns the remaining data as a tibble. By creating this helper, I made sure that I would not have to repeat the same API-response processing everytime I made a request. 

```{r}
# establish API key

api_key <- "9565ad3d32410ea2b7caa601ff9c5c4467f789d7"

# try calling to API using example 

url1 <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums?",
  "get=SEX,PWGTP,MAR&",
  "for=state:*&",
  "SCHL=24&",
  "key=",
  api_key
)

response <- GET(url1)

response

# convert response to text/character string & 
# parse the JSON text into an R object
data <- fromJSON(rawToChar(response$content))

# look at returned data
head(data)

# make the first row the column names, 
# then remove the first row
column_names <- data[1, ]
data <- data[-1, ]

# convert to a tibble
data <- as_tibble(data)

# make sure tibble has the correct column names
names(data) <- column_names

# check
head(data)

# use process to create a helper function & then test it

query_api <- function(url) {
  
  response <- GET(url)
  
  data <- fromJSON(rawToChar(response$content))
  
  column_names <- data[1, ]
  data <- data[-1, ]
  
  data <- as_tibble(data)
  
  names(data) <- column_names
  
  return(data)
}

query_api_test <- query_api(url1)

query_api_test
```

## Parsing Variable Values

The next step was to obtain and understand the metadata for the variables used in the project. The Census API provides a JSON file for the variables of each survey year, which contains various information about the variables. 

I created a list of the PUMS variables required for the project, `PUMS_var`, and then created the `get_var()` function to retrieve the metadata. Since the Census API uses `ST` for state in 2021 and 2022 but `STATE` in 2023 and 2024, I had to account for that in the function so that `get_var()` could be used for all four years. 

```{r}
# check 2024 PUMS variable data

url_2024 <- "https://api.census.gov/data/2024/acs/acs1/pums/variables.json"

info_2024 <- GET(url_2024)

data_2024 <- fromJSON(rawToChar(info_2024$content))

# inspect structure & variables 

str(data_2024, max.level = 2)
names(data_2024)
names(data_2024$variables) [1:10]

# create function & test it

PUMS_var <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP",
  "FER", "HHL", "SCH", "SCHL", "SEX",
  "REGION", "DIVISION")

get_var <- function(year) {
  
  url1 <- paste0(
    "https://api.census.gov/data/",
    year,
    "/acs/acs1/pums/variables.json"
  )
  
  response <- GET(url1)
  
  data <- fromJSON(rawToChar(response$content))
  
  # use ST for 2021 and 2022
  # use STATE for 2023 and 2024
  var <- PUMS_var
  if (year %in% c(2021, 2022)) {
    var <- c(var, "ST")
  } else {
    var <- c(var, "STATE")
  }
  
  variables <- data$variables[var]
  
  return(variables)
}

test_2024 <- get_var(2024)

names(test_2024)
```

The function successfully returned the requested variables for 2024, so I repeated the process for 2021 through 2023. I saved the metadata for each survey year as an `.rds` file in the `data` folder. Saving the metadata locally will allow for other functions to use the same information.

```{r}
data_2024 <- get_var(2024)
names(data_2024)

data_2023 <- get_var(2023)
names(data_2023)

data_2022 <- get_var(2022)
names(data_2022)

data_2021 <- get_var(2021)
names(data_2021)

# save as .rds in data folder

saveRDS(data_2024, "data/data_2024.rds")
saveRDS(data_2023, "data/data_2023.rds")
saveRDS(data_2022, "data/data_2022.rds")
saveRDS(data_2021, "data/data_2021.rds")
```

### PUMS Variable Reference

Since the Census variable names are abbreviations, I also created a reference table that showed the variables' names and descriptions. The descriptions come from the labels in the Census metadata. This will make the variables and results easier to interpret without having to repeatedly look up what each variable means. 

```{r}
var_ref <- tibble(
  Variable = PUMS_var,
  Description = sapply(PUMS_var, function(x) {
    data_2024[[x]]$label
  })
)
var_ref
```

## Parsing API Data

I then created helper functions for processing the returned data from the API. I created these helper functions as I was creating the main function (`query_census()`), but placed them before the main function. These helper functions were for two parts of the main function that required a little bit more work and therefore would keep the the main function looking neater. 

### Helper Functions

The first helper function I created was `get_geo()`, which would allow me to retrieve the appropriate geographic values from the metadata. It lets the main function accept a geographic name and determine the corresponding Census code. It also handles the difference between `ST` and `STATE` across the survey years and removes Puerto Rico as an acceptable value with `!grepl()`. Since `grepl()`  returns `TRUE` when the text is found and `FALSE` when it is not, and I wanted to exclude Puerto Rico from the data, I put `!` before `grepl()` to reverse the results. 

```{r}
# make helper function to return available geography 

get_geo <- function(year, geography) {
  
  # load variable values for selected year
  var_data <- readRDS(
    paste0(
      "data/data_",
      year,
      ".rds"
    )
  )
  
  # use ST for 2021/2022 and STATE for 2023/2024
  if (geography == "State" && year %in% c(2021, 2022)) {
    geography <- "ST"
  } else {
    geography <- toupper(geography)
  }
  
  # get selected geography & match them
  geo_data <- var_data[[geography]] 
  
  # return geography codes and names
  geo_val <- unlist(geo_data$values$item)
  
  # remove Puerto Rico 
  geo_val <- geo_val[!grepl("Puerto Rico", geo_val)
                     ]
  
  return(geo_val)
}

# test helper function
get_geo(2024, "State")
get_geo(2023, "Region")
get_geo(2022, "Division")
```

The second helper function I created was `parse_time()`, which processes `JWAP` and  `JWDP`. I had to use AI/LLMs to help figure this part out, but I think it may have become more complicated than it needed to be.

I first converted the codes to character values so they could be matched to the names in the metadata. The code `"0"` represents a missing value, so I changed those values to return as `NA`. I then used `sprintf()` to add leading zeroes to the codes so that they matched the three-digit codes used in the metadata. 

Then I used the codes to look up the corresponding time intervals. `sub()` separates the beginning and ending times of each interval. I converted those strings into time values with `strptime()`, then calculated the midpoint by adding half of the difference between the ending and beginning times to the starting time. I then used `hms::as_hms()` to keep only the time of day, so that the returned value represents the approximate midpoint of the five-minute interval rather than the original Census code. 

```{r}
# make helper function that parses the time strings to return midpoint as time value

parse_time <- function(x, time_values) {
  
  # convert codes to character 
  x <- as.character(x)
  
  # return missing time values ("0") as NA
  x[x == "0"] <- NA
  
  # add leading 0's so codes match metadata
  x <- ifelse(
    is.na(x), 
    NA, 
    sprintf("%03d", as.numeric(x)
            )
    )
  
  # find time interval for each code
  intervals <- unname(time_values[x])
  
  # get beginning and ending times
  start <- sub(" to .*", "", intervals)
  end <- sub(".* to ", "", intervals)
  
  # change Census AM/PM to standard AM/PM
  start <- gsub("a\\.m\\.", "AM", start)
  start <- gsub("p\\.m\\.", "PM", start)
  end <- gsub("a\\.m\\.", "AM", end)
  end <- gsub("p\\.m\\.", "PM", end)
  
  # convert into time values
  start_time <- strptime(start, format = "%I:%M %p")
  end_time <- strptime(end, format = "%I:%M %p")
  
  # calculate midpoint of each interval
  midpoint <- start_time + (end_time - start_time) / 2
  
  # return only time of day
  midpoint <- hms::as_hms(midpoint)
  
  return(midpoint)
}
```

### Main Function

I created the main function, `query_census()`, with the purpose of making it easier to request PUMS data from the API while also handling validation and data processing automatically. 

The function allows the user to specify the survey year, numeric variables, categorical variables, a geographic level, and a specific geographic area. I included several input checks that provide error messages when invalid inputs are supplied. The function makes sure that the year is one of the four years included in this project, that the requested variables are allowed, that `PWGTP` is included, and that the geographic areas are valid. It also prevents Puerto Rico from being requested. 
The function uses the metadata stored in the .rds files to identify geographic codes. This allows the user to enter a geographic name instead of having to know the Census code for that state. The user can also enter `all` to request all values at the selected geographic level.

After retrieving the data from the embedded API query, the function converts the numeric variables to numeric values and categorical variables to factors using the labels stored in the metadata. The two time variables, `JWAP` and `JWDP`, are processed using `parse_time()`. I also add the `census` class to the returned tibble so that the summary and plot functions can use the custom methods created later. 

```{r}
# create main parsing function 

# first set up description and defaults 

query_census <- function(
    year = 2024,
    numeric_vars = c("AGEP", "PWGTP"),
    categorical_vars = "SEX",
    geography = "State",
    specific_geography = NULL,
    api_key
) 
  {
  # check for valid year
  if (!year %in% c(2021, 2022, 2023, 2024)) {
    stop("year must be in 2021, 2022, 2023, or 2024")
  }
  
  # specify numeric variables that can be returned
  allowed_num <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP")
  
  # check
  if (!all(numeric_vars %in% allowed_num)) {
    stop("Invalid numeric variable. Choose from AGEP, GASP, GRPIP, JWAP, JWDP, JWMNP, and PWGTP.")
  }
  
  # make sure PWGTP is always returned 
  if (!"PWGTP" %in% numeric_vars) {
    stop("PWGTP must be included.")
  }
  
  # make sure another numeric variable is returned 
  if (length(numeric_vars) < 2) {
    stop("numeric_vars must included PWGTP and at least one other numeric variable.")
  }
  
  # specify categorical variables that can be returned 
  allowed_cat <- c("FER", "HHL", "SCH", "SCHL", "SEX")
  
  # check
  if (!all(categorical_vars %in% allowed_cat)) {
    stop("Invalid categorical variable. Choose from FER, HHL, SCH, SCHL, and SEX.")
  }
  
  # make sure at least one categorical variable is returned 
  if (length(categorical_vars) < 1) {
    stop("At least one categorical variable must be included.")
  }
  
  # specify geography levels that can be returned
  allowed_geo <- c("State", "Region", "Division")
  
  # make sure only one level can be returned 
  if (length(geography) != 1 || !geography %in% allowed_geo) {
    stop("geography must be either State, Region, or Division.")
  }
  
  # specify the available geography that can be subset 
  geo_values <- get_geo(year, geography)
  
  # if no specific_geography, use North Carolina as default 
  if(is.null(specific_geography)) {
    specific_geography <- "North Carolina"
  }
  
  # return error if user requests Puerto Rico
  if (grepl("Puerto Rico", specific_geography, ignore.case = TRUE)) {
    stop("Puerto Rico is not included in this function.")
  }

  # find Census code for selected geography
  if (specific_geography == "all") {
    geo_code <- NULL
  } else {
    match <- grepl(
      specific_geography,
      geo_values,
      fixed = TRUE
    )
    geo_code <- names(geo_values)[match]
  }
  
  # make sure valid specific geography is returned 
  if (length(geo_code) == 0 && specific_geography != "all") {
    stop("Invalid specific_geography.")
  }
  
  # create API query
  get_variables <- paste(c(numeric_vars, categorical_vars),
                         collapse = ","
                         )
  
  if (specific_geography == "all") {
    for_part <- paste0(
      tolower(geography),
      ":*"
    )
  } else {
    for_part <- paste0(
      tolower(geography),
      ":",
      geo_code
    )
  }
  
  # create API URL
  url2 <- paste0(
    "https://api.census.gov/data/",
    year,
    "/acs/acs1/pums?",
    "get=",
    get_variables,
    "&for=",
    for_part,
    "&key=",
    api_key
  )
  
  # query API
  data <- query_api(url2)
  
  # convert numeric variables
  for (var in numeric_vars) {
    if (var %in% c("JWAP", "JWDP")) {
      
      # get metadata
      time_url <- paste0(
        "https://api.census.gov/data/",
        year,
        "/acs/acs1/pums/variables/",
        var,
        ".json"
        )
      
      time_response <- GET(time_url)
      
      time_metadata <- fromJSON(
        rawToChar(time_response$content)
        )
      
      # get time codes and their labels
      time_values <- unlist(
        time_metadata$values$item
      )
      
      # convert time codes to midpoint times
      data[[var]] <- parse_time(
        data[[var]],
        time_values
      )
    } else {
      
      # convert regular numeric variables to numeric
      data[[var]] <- as.numeric(data[[var]])
    }
  }
  
  # load variable metadata
  var_data <- readRDS(paste0("data/data_", year, ".rds"))
  
  # convert categorical variables to factors
  for (var in categorical_vars) {
    value_labels <- unlist(var_data[[var]]$values$item)
    
    data[[var]] <- factor(
      data[[var]],
      levels = names(value_labels),
      labels = value_labels
    )
  }
  
  # give the data the census class
  class(data) <- c("census", class(data))
  
  return(data)
}
```

Because `JWAP` and `JWDP` are represented by coded five-minute intervals, I examined their data separately to understand how the codes correspond to actual times. I retrieved the metadata and saw that it provides the time intervals associated with each code. I used those with `parse_time()` to convert the values into the midpoint times. I then used `query_census()` to verify that returned variables' values were converted to time values. 

```{r}
# JWAP metadata
url <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums/variables/JWAP.json"
)

response <- GET(url)

jwap_metadata <- fromJSON(rawToChar(response$content))

jwap_values <- unlist(jwap_metadata$values$item)

head(jwap_values)

arrive_test <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
head(arrive_test$JWAP)

jwap_time <- parse_time(
  arrive_test$JWAP,
  jwap_values
)
head(jwap_time)
```

```{r}
# JWDP metadata 

url <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums/variables/JWDP.json"
)

response_jwdp <- GET(url)

jwdp_metadata <- fromJSON(rawToChar(response_jwdp$content))

jwdp_values <- unlist(jwdp_metadata$values$item)

head(jwdp_values)

depart_test <- query_census(
  numeric_vars = c("JWDP", "PWGTP"),
  api_key = api_key
)

jwdp_time <- parse_time(
  depart_test$JWDP,
  jwdp_values
)

head(jwdp_time)
```

### Function Tests

I tested `query_census()` using both valid and invalid inputs. The tests include different years, numeric and categorical variables, geographic levels, specific geographic areas, and special time variables. I also tested the default arguments. 

For the invalid tests, I used years, variables, and geographic values that would not be allowed to be requested. This showed that the function not only works successfully, but also provides an error when the user provides an invalid argument.

I also checked the classes and factor levels of the returned variables to make sure that the categorical variables are factors, the time variables are stored as time values, and that the overall tibble had the `census` class. 

```{r}
# test years

# valid
test_2021 <- query_census(
  year = 2021,
  api_key = api_key
  )
head(test_2021)

test_2023 <- query_census(
  year = 2023,
  api_key = api_key
  )
head(test_2023)

# invalid
query_census(
  year = 2002,
  api_key = api_key
  )

# test numeric variables

# valid
test_numeric <- query_census(
  numeric_vars = c("AGEP", "PWGTP"),
  api_key = api_key
  )
head(test_numeric)

# invalid numeric variables
query_census(
  numeric_vars = c("BDSP", "PWGTP"),
  api_key = api_key
  )

query_census(
  numeric_vars = c("AGEP", "GASP"),
  api_key = api_key
  )

query_census(
  numeric_vars = "PWGTP",
  api_key = api_key
  )

# test categorical variables

# valid
test_categorical <- query_census(
  numeric_vars = c("GASP", "PWGTP"),
  categorical_vars = c("FER", "SEX"),
  api_key = api_key
  )

head(test_categorical)
class(test_categorical$SEX)
levels(test_categorical$SEX)

# invalid
query_census(
  categorical_vars = "WAOB",
  api_key = api_key
  )

# test geography

# valid 
test_geo <- query_census(
  geography = "Region",
  specific_geography = "South",
  api_key = api_key
  )
head(test_geo)

test_geo2 <- query_census(
  specific_geography = "California",
  api_key = api_key
  )
head(test_geo2)

# invalid
query_census(
  geography = "All",
  api_key = api_key
  )

query_census(
  geography = c("Region", "Division"),
  api_key = api_key
  )

query_census(
  specific_geography = "Puerto Rico",
  api_key = api_key
  )

query_census(
  specific_geography = "Cuba",
  api_key = api_key
  )

# test default query
test_default <- query_census(
  api_key = api_key
  )
head(test_default)

# test time variable
test_time <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
head(test_time$JWAP)
class(test_time$JWAP)

# test census class
class(test_time)

# general test
gen_test <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
class(gen_test$JWAP)
class(gen_test$SEX)
levels(gen_test$SEX)
head(gen_test$SEX)
```

## Combining Years

I then created `query_census_years()` so that multiple survey years could be retrieved and combined within one call, instead of just a single year being retrieved. The function calls `query_census()` once for every requested year and combines the resulting tibbles using `bind_rows()`; this is so I did not have to rewrite the API query for each year. 

I also added a `year` variable to each year's data before combining them so that the same variables could be compared across multiple survey years. I gave the combined data the `census` class so that R knows to use my custom summary and plot functions later. 

```{r}
# create function to specify multiple years of survey data

query_census_years <- function(
    years = c(2021, 2022, 2023, 2024),
    numeric_vars = c("AGEP", "PWGTP"),
    categorical_vars = "SEX",
    geography = "State",
    specific_geography = NULL,
    api_key
) {
  
  # check for valid years
  if (!all(years %in% c(2021, 2022, 2023, 2024))) {
    stop("years must be in 2021, 2022, 2023, or 2024")
  }
  
  # make sure at least one year in specified 
  if (length(years) < 1) {
    stop("At least one year must be specified.")
  }

  # create a list to store each year's data
  data_list <- list()

  # query each year
  for (year in years) {
    data <- query_census(
      year = year,
      numeric_vars = numeric_vars,
      categorical_vars = categorical_vars,
      geography = geography,
      specific_geography = specific_geography,
      api_key = api_key
    )
    
    # add year variable
    data$year <- year
    
    data_list[[as.character(year)]] <- data
  }
  
  # combine all years into one tibble
  data <- bind_rows(data_list)
  
  # give census class
  class(data) <- c("census", class(data))
  
  return(data)
}
```

I then tested the function with different combinations of years and variables. I also tests the `all` option to retrieve data from all states. 

These tests showed that the function works properly and can be reused with different inputs. 
```{r}
# test function

combined_test <- query_census_years(
  years = c(2022, 2024),
  api_key = api_key
)
head(combined_test)
table(combined_test$year)
class(combined_test)

combined_all_years <- query_census_years(
  years = c(2021, 2022, 2023, 2024),
  api_key = api_key
)
head(combined_all_years)
table(combined_all_years$year)

combined_test2 <- query_census_years(
  years = c(2022, 2023, 2024),
  numeric_vars = c("GASP", "PWGTP"),
  categorical_vars = c("FER", "SEX"),
  geography = "State",
  specific_geography = "California",
  api_key = api_key
)
head(combined_test2)

combined_all_states <- query_census_years(
  years = c(2023, 2024),
  specific_geography = "all",
  api_key = api_key
)
table(combined_all_states$year)
```

# Part 2: Writing a Generic Function for Summarizing

Part 2 of the project focuses on creating numerical and graphical summary functions using the `census` class.  

### Numeric Summary: Weighted Summary Function

Because `PWGTP` represents the number of people represented by each record, a regular mean would give every row the same influence, even though different rows can represent different numbers of people. Therefore, I created a summary function that would calculate weighted summaries, `summary.census()`. 

For numeric variables, the function calculates a weighted mean and weighted standard deviation using the person weights. I created `keep` to make sure missing values did not affect the calculations. For categorical variables, it calculates weighted counts for each category, also using person weights. `year` is treated as categorical when multiple years are requested, since I want to compare the distribution of observations across survey years.

The function lets users specify which numeric and categorical variables to summarize, while excluding the person weight, `PWGTP`, and the time variables, `JWAP` and `JWDP` since they are time values. 

```{r}
# create a function that 
# produces means & sds for numeric variables & 
# counts for categorical variables 

summary.census <- function(
    object,
    numeric_vars = NULL,
    categorical_vars = NULL
) {
  # default numeric variables 
  if (is.null(numeric_vars)) {
    numeric_vars <- names(object) [
      sapply(object, is.numeric)
    ]
    
    numeric_vars <- numeric_vars[!numeric_vars %in% c("PWGTP", "JWAP", "JWDP", "year")]
  }
  
  # default categorical variables
  if (is.null(categorical_vars)) {
    categorical_vars <- names(object) [
      sapply(object, is.factor)
    ]
    
    # include year as a categorical variable
    if ("year" %in% names(object)) {
      categorical_vars <- unique(
        c(categorical_vars, "year")
      )
    }
  }
  
  # create list for numeric summaries
  numeric_summary <- list()
  
  # calculate weighted mean and standard deviation
  for (var in numeric_vars) {
    x <- object[[var]]
    w <- object$PWGTP
    
    keep <- !is.na(x) & !is.na(w)
    x <- x[keep]
    w <- w[keep]
    
    weighted_mean <- sum(x * w) / sum(w)
    
    weighted_sd <- sqrt(
      sum(x^2 * w) / sum(w) -
        weighted_mean^2
    )
    
    numeric_summary[[var]] <- c(
      mean = weighted_mean,
      sd = weighted_sd
    )
  }
  
  # create list for categorical summaries
  categorical_summary <- list()
  
  # calculate weighted counts
  for (var in categorical_vars) {
    x <- object[[var]]
    
    # add person weights within each category
    categorical_summary[[var]] <- tapply(
      object$PWGTP,
      x,
      sum,
      na.rm = TRUE
    )
  }
  
  # return named lists
  return(
    list(
      numeric = numeric_summary,
      categorical = categorical_summary
    )
  )
}
```

I tested `summary_census()` to show that the function can summarize different combinations of data while using `PWGTP` to account for the survey weights. I first tested the default arguments so the function automatically identify the appropriate numeric and categorical variables. I then specified individual variables and combination of variables. I also tested it to ensure year is evaluated as a categorical variable. 

```{r}
# test function

summary(combined_test)

summary(
  combined_test,
  numeric_vars = "AGEP",
  categorical_vars = c("SEX", "year")
)

summary(
  combined_test,
  numeric_vars = c("AGEP", "GRPIP"),
  categorical_vars = "SEX"
)
```

### Graphical Summary: Plotting Function

I then created `plot_census()` to produce weighted boxplots for the returned Census data. It requires the user to specify a categorical variable and a numeric variable, and uses `PWGTP` as the weight so the plot accounts for the fact that each PUMS record may represent multiple people. 

I also made sure `year` is treated as a categorical variable when used for the x-axis, so the boxplots show separate distributions for each survey year. 

```{r}
# create a generic plot() function that 
# creates a boxplot of the numeric variable 
# at the levels of the categorical variables 

plot.census <- function(
    object,
    categorical_var,
    numeric_var
) {
  
  # make year a categorical variable 
  if (categorical_var == "year") {
    object$year <- factor(object$year)
  }
  
  ggplot(
    object,
    aes(
      x = get(categorical_var),
      y = get(numeric_var),
      weight = PWGTP
    )
  ) +
    geom_boxplot(width = 0.6) +
    labs(
      x = categorical_var,
      y = numeric_var,
      title = paste(numeric_var, "by", categorical_var)
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.title = element_text(face = "bold")
    )
}
```

I tested `plot.census()` using different categorical and numeric variables. The first example compares age across sex, while the second compares age across survey years. By using different inputs, the function shows that it can be reused with different combinations of variables. 

```{r}
# test plot function

plot(
  combined_test,
  categorical_var = "SEX",
  numeric_var = "AGEP"
)

plot(
  combined_test,
  categorical_var = "year",
  numeric_var = "AGEP"
)
```

# Investigation

For my investigation, I wanted to examine educational attainment in North Carolina from  2021 to 2024. I was interested in how educational attainment is associated with age and whether the distribution differs between males and females. 

## Educational Attainment Across Survey Years

I began by retrieving the data using my `query_census_years()` function. 

```{r}
# get data

nc_edu <- query_census_years(
  years = c(2021, 2022, 2023, 2024),
  numeric_vars = c("AGEP", "PWGTP"),
  categorical_vars = c("SCHL", "SEX"),
  geography = "State",
  specific_geography = "North Carolina",
  api_key = api_key
)

head(nc_edu)
table(nc_edu$year)
```

I used my `summary()` method to examine educational attainment separately for the 2021 and 2024 survey years. 

```{r}
# investigate educational attainment across years 

summary(
  nc_edu[nc_edu$year == 2021, ],
  numeric_vars = "AGEP",
  categorical_vars = c("SCHL", "year")
)

summary(
  nc_edu[nc_edu$year == 2024, ],
  numeric_vars = "AGEP",
  categorical_vars = c("SCHL", "year")
)
```

The weighted results show some changes in educational attainment across the four years. The weighted number of people with a bachelor's degree increased from about 1.69 million in 2021 to 1.88 million in 2024. The weighted number with a master's degree also increased, from about 713,000 in 2021 to 840,000 in 2024. The weighted number with a doctorate degree increased from about 112,00 to 142,000. 

## Educational Attainment and Age

Next, I examined the relationship betwen educational attainment and age. While I used my `plot.census()` function to create a weighted boxplot of age for each educational attainment category, the large number of `SCHL` categories made the plot difficult to read. Instead, I used a horizontal weighted boxplot to compare the age distributions within each educational attainment category. The Census data contains many educational attainment categories, so placing the categories on the y-axis made the longer category names easier to read and compare.  

```{r}
# investigate education and age

plot(
  nc_edu,
  categorical_var = "SCHL",
  numeric_var = "AGEP"
)

ggplot(
  nc_edu,
  aes(
    y = SCHL,
    x = AGEP,
    weight = PWGTP
  )
) +
  geom_boxplot() +
  labs(
    y = "Educational Attainment",
    x = "Age",
    title = "Age by Educational Attainment"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold")
  )
```

The boxplot shows that age distributions vary across the educational attainment categories. The median ages for several of the college and graduate-level categories are around the mid-40s to early 50s. The plot also shows that several categories have a wide range of ages, indicating that people within the same educational attainment category can vary considerably in age.

## Educational Attainment and Sex

Then I investigated whether educational attainment differs between males and females. I separated the data into male and female observations, then usedd my weighted `summary()` function to compare the educational attainment categories for the two groups.

```{r}
# investigate education and sex

male_data <- nc_edu[
  nc_edu$SEX == "Male",
]

female_data <- nc_edu[
  nc_edu$SEX == "Female",
]

summary(
  male_data,
  numeric_vars = "AGEP",
  categorical_vars = "SCHL"
)

summary(
  female_data,
  numeric_vars = "AGEP",
  categorical_vars = "SCHL"
)
```

The weighted summaries show differences between males and females in multiple educational attainment categories. The weighted number of people with a bachelor's degree was about 3.30 million among males and 3.91 million among females. The difference was also noticeable for master's degrees, with about 1.26 million males and 1.79 million females.

## Sex and Age

Because age may also be related to educational attainment, I compared the age distributions of males and females.

```{r}
# compare age distributions by sex

plot(
  nc_edu,
  categorical_var = "SEX",
  numeric_var = "AGEP"
)
```

The boxplot shows that the age distributions of males and females are similar, although females have a slightly higher median age. The median age for males is about 37 years and 40 years for females. 

## Conclusion

Overall, this investigation showed how my Census API PUMS functions can be used to examine questions. It also showed that some variables, such as `SCHL`, may require additional grouping before they can be easily interpreted.

In North Carolina, the  weighted counts for bachelor's, master's, and doctorate degrees increased between 2021 and 2024. The boxplot comparing age and educational attainment showed that age distributions varied across education categories. However, because `SCHL` contains many categories, the comparison was somewhat difficult to interpret in detail. Grouping some of the educational attainment categories could make this comparison easier to understand. 

The weighted counts also showed differences between males and females across several educational attainment categories. Females had higher weighted counts for bachelor's and master's degrees. The age distributions of males and females were relatively similar, although females had a slightly higher median age.

These results show associations in the data instead of relationships. Since the educational attainment comparisons were based on weighted counts, using percentages would provide a more direct way to compare the distributions across years and between males and females. 