---
title: "Project 1"
author: "Maya Szymborski"
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: right
editor: source
execute:
  error: false
  warning: false
  message: false
---

# Introduction

The goal of this project is to create reusable, generic functions for working with data from the U.S. Census Bureau's Public Use Microdata Sample (PUMS) API. The functions should be able to retrieve data from different survey years, variables, and geographic areas, and return the data consistently without manually repeating the same data-processing steps.  

The PUMS API provides psuedo person-level records, but the records do not represent exactly one person but can represent multiple people. The `PWGTP` variable is a person weight, which indicates how many people a particular record represents. Because of this, the functions need to account for these weights. 

I focused on the PUMS data from 2021 through 2024, and excluded Puerto Rico as stated in the project instructions. My approach was to first understand how the Census API returned data, then create helper functions to process the metadata and special variables, and finally build and test functions that could retrieve, combine, summarize, and plot the data. 

# Part 1: Data Processing

Part 1 of the project focuses on retrieving and processing the Census API's PUMS data.

```{r}
# attach packages 
library(httr)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(quantreg)
```

## Explore PUMS Census API

I began by making a request to the 2024 Census PUMS API using the example provided in the project instructions. This let me see what the returned data from the Census API looked like. 

Since the response was returned as a JSON, I converted it to an R object. The first row contained the variable names while the rest of the rows contained the observations. I separated the column names from the observations, and then converted the response to a tibble. 

I then used those steps to create a helper function, `query_api`. This function takes an inputted URL, sends the request, converts the JSON response into an R object, uses the first row as the column names, and returns the reamining data as a tibble. By creating this helper, I made sure that I would not have to repeat the same API-response processing everytime I made a request. 

```{r}
# establish API key

api_key <- "9565ad3d32410ea2b7caa601ff9c5c4467f789d7"

# try calling to API using example 

url1 <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums?",
  "get=SEX,PWGTP,MAR&",
  "for=state:*&",
  "SCHL=24&",
  "key=",
  api_key
)

response <- GET(url1)

response

# convert response to text/character string & 
# parse the JSON text into an R object
data <- fromJSON(rawToChar(response$content))

# look at returned data
head(data)

# make the first row the column names, 
# then remove the first row
column_names <- data[1, ]
data <- data[-1, ]

# convert to a tibble
data <- as_tibble(data)

# make sure tibble has the correct column names
names(data) <- column_names

# check
head(data)

# use process to create a helper function & then test it

query_api <- function(url) {
  
  response <- GET(url)
  
  data <- fromJSON(rawToChar(response$content))
  
  column_names <- data[1, ]
  data <- data[-1, ]
  
  data <- as_tibble(data)
  
  names(data) <- column_names
  
  return(data)
}

query_api_test <- query_api(url1)

query_api_test
```

## Parsing Variable Values

The next step was to obtain and understand the metadata for the variables used in the project. The Census API provides a JSON file for the variables of each survey year, which contains various information about the variables. 

I created a list of the PUMS variables required for the project, `PUMS_var`, and then created the `get_var()` function to retrieve the metadata. Since the Census API uses `ST` for state in 2021 and 2022 but `STATE` in 2023 and 2024, I had to account for that in the function so that `get_var()` could be used for all four years. 

```{r}
# check 2024 PUMS variable data

url_2024 <- "https://api.census.gov/data/2024/acs/acs1/pums/variables.json"

info_2024 <- GET(url_2024)

data_2024 <- fromJSON(rawToChar(info_2024$content))

# inspect structure & variables 

str(data_2024, max.level = 2)
names(data_2024)
names(data_2024$variables) [1:10]

# create function & test it

PUMS_var <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP",
  "FER", "HHL", "SCH", "SCHL", "SEX",
  "REGION", "DIVISION")

get_var <- function(year) {
  
  url1 <- paste0(
    "https://api.census.gov/data/",
    year,
    "/acs/acs1/pums/variables.json"
  )
  
  response <- GET(url1)
  
  data <- fromJSON(rawToChar(response$content))
  
  # keep requested variables
  var <- PUMS_var
  if (year == 2021 | year == 2022) {
    var <- c(var, "ST")
  } else {
    var <- c(var, "STATE")
  }
  
  variables <- data$variables[var]
  
  return(variables)
}

test_2024 <- get_var(2024)

names(test_2024)
```

The function successfully returned the requested variables for 2024, so I repeated the process for 2021 through 2023. I saved the metadata for each survey year as an `.rds` file in the `data` folder. Saving the metadata locally will allow for other functions to use the same information.

```{r}
data_2024 <- get_var(2024)
names(data_2024)

data_2023 <- get_var(2023)
names(data_2023)

data_2022 <- get_var(2022)
names(data_2022)

data_2021 <- get_var(2021)
names(data_2021)

# save as .rds in data folder

saveRDS(data_2024, "data/data_2024.rds")
saveRDS(data_2023, "data/data_2023.rds")
saveRDS(data_2022, "data/data_2022.rds")
saveRDS(data_2021, "data/data_2021.rds")
```

### PUMS Variable Reference

Since the Census variable names are abbreviations, I also created a reference table that showed the variables' names and descriptions. The descriptions come from the labels in the Census metadata. This will make the variables and results easier to interpret without having to repeatedly look up what each variable means. 

```{r}
var_ref <- tibble(
  Variable = PUMS_var,
  Description = sapply(PUMS_var, function(x) {
    data_2024[[x]]$label
  })
)
var_ref
```

## Parsing API Data

I then created helper functions for processing the returned data from the API. I created these helper functions as I was creating the main function (`query_census()`), but placed them before the main function. These helper functions were for two parts of the main function that required a little bit more work and therefore would keep the the main function looking neater. 

### Helper Functions

The first helper function I created was `get_geo()`, which would allow me to retrieve the appropriate geographic values from the metadata. It lets the main function accept a geographic name and determine the corresponding Census code. It also handles the difference between `ST` and `STATE` across the survey years and removes Puerto Rico as an acceptable value. 

```{r}
# make helper function to return available geography 

# inspect for specific_geography
data_2024$STATE$values
data_2024$REGION$values
data_2024$DIVISION$values

state_values <- data_2024$STATE$values$item
state_values
names(state_values)
unlist(state_values) # turn nested list into a simplified list

get_geo <- function(year, geography) {
  
  # load variable values for selected year
  var_data <- readRDS(
    paste0(
      "data/data_",
      year,
      ".rds"
    )
  )
  
  # use ST for 2021/2022 and STATE for 2023/2024
  if (geography == "State" && year %in% c(2021, 2022)) {
    geography <- "ST"
  } else {
    geography <- toupper(geography)
  }
  
  # get selected geography & match them
  geo_data <- var_data[[geography]] 
  
  # return geography codes and names
  geo_val <- unlist(geo_data$values$item)
  
  # remove Puerto Rico using `!grepl` which means "does not contain"
  geo_val <- geo_val[!grepl("Puerto Rico", geo_val)
                     ]
  
  return(geo_val)
}

# test helper function
get_geo(2024, "State")
get_geo(2023, "Region")
get_geo(2022, "Division")
```

The second helper function I created was `parse_time()`, which would process `JWAP` and  `JWDP`. These variables are provided as codes that represent five-minute time intervals. I converted the codes to their corresponding intervals, calculated the midpoint of each interval, and then returned the result as a time value. Missing time values are returned as `NA`. 

```{r}
# make helper function that parses the time strings to return midpoint as time value

parse_time <- function(x, time_values) {
  
  # convert values to character to match metadata names
  x <- as.character(x)
  
  # return missing time values as NA
  x[x == "0"] <- NA
  
  # add leading 0's so codes match Census metadata
  x <- ifelse(
    is.na(x), 
    NA, 
    sprintf("%03d", as.numeric(x)
            )
    )
  
  # get time interval for each code
  intervals <- unname(time_values[x])
  
  # extract beginning and ending times
  start <- sub(" to .*", "", intervals)
  end <- sub(".* to ", "", intervals)
  
  # change Census AM/PM to standard AM/PM
  start <- gsub("a\\.m\\.", "AM", start)
  start <- gsub("p\\.m\\.", "PM", start)
  end <- gsub("a\\.m\\.", "AM", end)
  end <- gsub("p\\.m\\.", "PM", end)
  
  # convert times to POSIXct
  start_time <- strptime(start, format = "%I:%M %p")
  end_time <- strptime(end, format = "%I:%M %p")
  
  # calculate midpoint of each interval
  midpoint <- start_time + (end_time - start_time) / 2
  
  # return only time of day
  midpoint <- hms::as_hms(midpoint)
  
  return(midpoint)
}
```

### Main Function

I created the main function, `query_census()`, with the purpose of making it easier to request PUMS data from the API while also handling validation and data processing automatically. 

The function allows the user to specify the survey year, numeric variables, categorical variables, a geographic level, and a specific geographic area. I included several input checks that provided error messages when invalid inputs are supplied. The function makes sure that the year is one of the four years included in this project, that the requested variables are allowed, that `PWGTP` is included, and that the geographic areas are valid. It also prevents Puerto Rico from being requested. 

The function uses the metadata stored in the .rds files to identify geographic codes. This allows the user to enter a geographic name instead of having to know the Census code for that state. The user can also enter `all` to request all values at the selected geographic level.

After retrieving the data from the embedded API query, the function converts the numeric variables to numeric values and categorical variables to factors using the labels stored in the metadata. The two times variables, `JWAP` and `JWDP, are processed using `parse_time()`. I also add the `census` class to the returned tibble so that the summary and plot functions can use the custom methods created later. 

```{r}
# create main parsing function 

# first set up description and defaults 

query_census <- function(
    year = 2024,
    numeric_vars = c("AGEP", "PWGTP"),
    categorical_vars = "SEX",
    geography = "State",
    specific_geography = NULL,
    api_key
) 
  {
  # check for valid year
  if (!year %in% c(2021, 2022, 2023, 2024)) {
    stop("year must be in 2021, 2022, 2023, or 2024")
  }
  
  # specify numeric variables that can be returned
  allowed_num <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "PWGTP")
  
  # check
  if (!all(numeric_vars %in% allowed_num)) {
    stop("Invalid numeric variable. Choose from AGEP, GASP, GRPIP, JWAP, JWDP, JWMNP, and PWGTP.")
  }
  
  # make sure PWGTP is always returned 
  if (!"PWGTP" %in% numeric_vars) {
    stop("PWGTP must be included.")
  }
  
  # make sure another numeric variable is returned 
  if (length(numeric_vars) < 2) {
    stop("numeric_vars must included PWGTP and at least one other numeric variable.")
  }
  
  # specify categorical variables that can be returned 
  allowed_cat <- c("FER", "HHL", "SCH", "SCHL", "SEX")
  
  # check
  if (!all(categorical_vars %in% allowed_cat)) {
    stop("Invalid categorical variable. Choose from FER, HHL, SCH, SCHL, and SEX.")
  }
  
  # make sure at least one categorical variable is returned 
  if (length(categorical_vars) < 1) {
    stop("At least one categorical variable must be included.")
  }
  
  # specify geography levels that can be returned
  allowed_geo <- c("State", "Region", "Division")
  
  # make sure only one level can be returned 
  if (length(geography) != 1 || !geography %in% allowed_geo) {
    stop("geography must be either Region, Division, or State.")
  }
  
  # specify the available geography that can be subset 
  geo_values <- get_geo(year, geography)
  
  # if no specific_geography, use North Carolina as default 
  if(is.null(specific_geography)) {
    specific_geography <- "North Carolina"
  }
  
  # return error if user requests Puerto Rico
  if (grepl("Puerto Rico", specific_geography, ignore.case = TRUE)) {
    stop("Puerto Rico is not included in this function.")
  }

  # find Census code for selected geography
  if (specific_geography == "all") {
    geo_code <- NULL
  } else {
    geo_code <- names(
    geo_values[grepl(
      specific_geography,
      geo_values,
      fixed = TRUE
    )]
  )
  }
  
  # make sure valid specific geography is returned 
  if (length(geo_code) == 0 && specific_geography != "all") {
    stop("Invalid specific_geography.")
  }
  
  # create API query
  get_variables <- paste(c(numeric_vars, categorical_vars),
                         collapse = ","
                         )
  
  if (specific_geography == "all") {
    for_part <- paste0(
      tolower(geography),
      ":*"
    )
  } else {
    for_part <- paste0(
      tolower(geography),
      ":",
      geo_code
    )
  }
  
  # create API URL
  url2 <- paste0(
    "https://api.census.gov/data/",
    year,
    "/acs/acs1/pums?",
    "get=",
    get_variables,
    "&for=",
    for_part,
    "&key=",
    api_key
  )
  
  # query API
  data <- query_api(url2)
  
  # convert numeric variables
  for (var in numeric_vars) {
    if (var %in% c("JWAP", "JWDP")) {
      
      # get metadata
      time_url <- paste0(
        "https://api.census.gov/data/",
        year,
        "/acs/acs1/pums/variables/",
        var,
        ".json"
        )
      
      time_response <- GET(time_url)
      
      time_metadata <- fromJSON(
        rawToChar(time_response$content)
        )
      
      time_values <- unlist(
        time_metadata$values$item
      )
      
      # convert Census time codes to midpoint times
      data[[var]] <- parse_time(
        data[[var]],
        time_values
      )
    } else {
      
      # convert regular numeric variables to numeric
      data[[var]] <- as.numeric(data[[var]])
    }
  }
  
  # load variable metadata
  var_data <- readRDS(paste0("data/data_", year, ".rds"))
  
  # convert categorical variables to factors
  for (var in categorical_vars) {
    levels <- unlist(var_data[[var]]$values$item)
    
    data[[var]] <- factor(
      data[[var]],
      levels = names(levels),
      labels = levels
    )
  }
  
  # give the data the census class
  class(data) <- c("census", class(data))
  
  return(data)
}
```

yada yada

```{r}
# JWAP metadata
url <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums/variables/JWAP.json"
)

response <- GET(url)

fromJSON(rawToChar(response$content))

jwap_metadata <- fromJSON(rawToChar(response$content))

jwap_values <- unlist(jwap_metadata$values$item)

head(jwap_values)

arrive_test <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
head(arrive_test$JWAP)

jwap_time <- parse_time(
  arrive_test$JWAP,
  jwap_values
)
head(jwap_time)
```

```{r}
# JWDP metadata 

url <- paste0(
  "https://api.census.gov/data/2024/acs/acs1/pums/variables/JWDP.json"
)

response_jwdp <- GET(url)

jwdp_metadata <- fromJSON(rawToChar(response_jwdp$content))

jwdp_values <- unlist(jwdp_metadata$values$item)

head(jwdp_values)

depart_test <- query_census(
  numeric_vars = c("JWDP", "PWGTP"),
  api_key = api_key
)

jwdp_time <- parse_time(
  depart_test$JWDP,
  jwdp_values
)

head(jwdp_time)
```

yada yada

### Function Tests

I tested `query_census()` using both valid and invalid inputs. The tests include different years, numeric and categorical variables, geographic levels, specific geographic areas, and special time variables. I also tested the default arguments. 

For the invalid tests, I used years, variables, and geographic values that would not be allowed to be requested. This showed that the function not only works successfully, but also provides an error when the user provides an invalid argument.

I also checked the classes and factor levels of the returned variables to make sure that the categorical variables are factors, the time variables are stored as time values, and that the overall tibble had the `census` class. 

```{r}
# test years

# valid
test_2021 <- query_census(
  year = 2021,
  api_key = api_key
  )
head(test_2021)

test_2023 <- query_census(
  year = 2023,
  api_key = api_key
  )
head(test_2023)

# invalid
query_census(
  year = 2002,
  api_key = api_key
  )

# test numeric variables

# valid
test_numeric <- query_census(
  numeric_vars = c("AGEP", "PWGTP"),
  api_key = api_key
  )
head(test_numeric)

# invalid numeric variables
query_census(
  numeric_vars = c("BDSP", "PWGTP"),
  api_key = api_key
  )

query_census(
  numeric_vars = c("AGEP", "GASP"),
  api_key = api_key
  )

query_census(
  numeric_vars = "PWGTP",
  api_key = api_key
  )

# test categorical variables

# valid
test_categorical <- query_census(
  numeric_vars = c("GASP", "PWGTP"),
  categorical_vars = c("FER", "SEX"),
  api_key = api_key
  )

head(test_categorical)
class(test_categorical$SEX)
levels(test_categorical$SEX)

# invalid
query_census(
  categorical_vars = "WAOB",
  api_key = api_key
  )

# test geography

# valid 
test_geo <- query_census(
  geography = "Region",
  specific_geography = "South",
  api_key = api_key
  )
head(test_geo)

test_geo2 <- query_census(
  specific_geography = "California",
  api_key = api_key
  )
head(test_geo2)

# invalid
query_census(
  geography = "All",
  api_key = api_key
  )

query_census(
  geography = c("Region", "Division"),
  api_key = api_key
  )

query_census(
  specific_geography = "Puerto Rico",
  api_key = api_key
  )

query_census(
  specific_geography = "Cuba",
  api_key = api_key
  )

# test default query
test_default <- query_census(
  api_key = api_key
  )
head(test_default)

# test time variable
test_time <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
head(test_time$JWAP)
class(test_time$JWAP)

# test census class
class(test_time)

# general test
gen_test <- query_census(
  numeric_vars = c("JWAP", "PWGTP"),
  api_key = api_key
)
class(gen_test$JWAP)
class(gen_test$SEX)
levels(gen_test$SEX)
head(gen_test$SEX)
```

## Combining Years

I then created `query_census_years()` so that multiple survey years could be retrieved and combined within one call, instead of just a single year being retrieved. The function calls `query_census()` once for every requested year and combines the resulting tibbles; this is so I did not have to rewrite the API query for each year. 

I also added a `year` variable to each year's data before combining them so that the same variables could be compared across multiple survey years. The resulting combined tibble retains the "census" class as well so that it can be passed directly to the custom summary and plotting functions later. 

```{r}
# create function to specify multiple years of survey data

query_census_years <- function(
    years = c(2021, 2022, 2023, 2024),
    numeric_vars = c("AGEP", "PWGTP"),
    categorical_vars = "SEX",
    geography = "State",
    specific_geography = NULL,
    api_key
) {
  
  # check for valid years
  if (!all(years %in% c(2021, 2022, 2023, 2024))) {
    stop("years must be in 2021, 2022, 2023, or 2024")
  }
  
  # make sure at least one year in specified 
  if (length(years) < 1) {
    stop("At least one year must be specified.")
  }

  # create a list to store each year's data
  data_list <- list()

  # query each year
  for (i in seq_along(years)) {
    data_list[[i]] <- query_census(
      year = years[i],
      numeric_vars = numeric_vars,
      categorical_vars = categorical_vars,
      geography = geography,
      specific_geography = specific_geography,
      api_key = api_key
    )
    
    # add year variable
    data_list[[i]]$year <- years[i]
  }
  
  # combine all years into one tibble
  data <- bind_rows(data_list)
  
  # give census class
  class(data) <- c("census", class(data))
  
  return(data)
}
```

I then tested the function with different combinations of years and variables. I also tests the `all` option to retrieve data from all states. 

These tests showed that the function works properly and can be reused with different inputs. 
```{r}
# test function

combined_test <- query_census_years(
  years = c(2022, 2024),
  api_key = api_key
)
head(combined_test)
table(combined_test$year)
class(combined_test)

combined_all_years <- query_census_years(
  years = c(2021, 2022, 2023, 2024),
  api_key = api_key
)
head(combined_all_years)
table(combined_all_years$year)

combined_test2 <- query_census_years(
  years = c(2022, 2023, 2024),
  numeric_vars = c("GASP", "PWGTP"),
  categorical_vars = c("FER", "SEX"),
  geography = "State",
  specific_geography = "California",
  api_key = api_key
)
head(combined_test2)

combined_all_states <- query_census_years(
  years = c(2023, 2024),
  specific_geography = "all",
  api_key = api_key
)
table(combined_all_states$year)
```

# Part 2: Writing a Generic Function for Summarizing

Part 2 of the project focuses on creating numerical and graphical summary functions using the `census` class.  

### Numeric Summary: Weighted Summary Function

Because `PWGTP` represents the number of people represented by each record, a regular mean would give every row the same influence, even though different rows can represent different numbers of people. Therefore, I created a summary function that would calculate weighted summaries, `summary.census()`. 

For numeric variables, the function calculates a weighted mean and weighted standard deviation using the person weights. For categorical variables, it calculates weighted counts for each category, also using person weights. `year` is treated as categorical when multiple years are requested, since I want to compare the distribution of observations across survey years. 
The function lets users specify which numeric and categorical variables to summarize, while excluding the person weight, `PWGTP`, and the time variables, `JWAP` and `JWDP` since they are time values. 

```{r}
# create a function that 
# produces means & sds for numeric variables & 
# counts for categorical variables 

summary.census <- function(
    object,
    numeric_vars = NULL,
    categorical_vars = NULL
) {
  # default numeric variables 
  if (is.null(numeric_vars)) {
    numeric_vars <- names(object) [
      sapply(object, is.numeric)
    ]
    
    numeric_vars <- setdiff(
      numeric_vars,
      c("PWGTP", "JWAP", "JWDP", "year")
    )
  }
  
  # default categorical variables
  if (is.null(categorical_vars)) {
    categorical_vars <- names(object) [
      sapply(object, is.factor)
    ]
    
    # include year as a categorical variable
    if ("year" %in% names(object)) {
      categorical_vars <- unique(
        c(categorical_vars, "year")
      )
    }
  }
  
  # create list for numeric summaries
  numeric_summary <- list()
  
  # calculate weighted mean and standard deviation
  for (var in numeric_vars) {
    x <- object[[var]]
    w <- object$PWGTP
    
    keep <- !is.na(x) & !is.na(w)
    x <- x[keep]
    w <- w[keep]
    
    weighted_mean <- sum(x * w) / sum(w)
    
    weighted_sd <- sqrt(
      sum(x^2 * w) / sum(w) -
        weighted_mean^2
    )
    
    numeric_summary[[var]] <- c(
      mean = weighted_mean,
      sd = weighted_sd
    )
  }
  
  # create list for categorical summaries
  categorical_summary <- list()
  
  # calculate weighted counts
  for (var in categorical_vars) {
    x <- object[[var]]
    
    categorical_summary[[var]] <- tapply(
      object$PWGTP,
      x,
      sum,
      na.rm = TRUE
    )
  }
  
  # return named lists
  return(
    list(
      numeric = numeric_summary,
      categorical = categorical_summary
    )
  )
}
```

I tested `summary_census()` to show that the function can summarize different combinations of data while using `PWGTP` to account for the survey weights. I first tested the default arguments so the function automatically identify the appropriate numeric and categorical variables. I then specified individual variables and combination of variables. I also tested it to ensure year is evaluated as a categorical variable. 

```{r}
# test function

summary(combined_test)

summary(
  combined_test,
  numeric_vars = "AGEP",
  categorical_vars = c("SEX", "year")
)

summary(
  combined_test,
  numeric_vars = c("AGEP", "GRPIP"),
  categorical_vars = "SEX"
)
```

### Graphical Summary: Plotting Function

I then created `plot_census()` to produce weighted boxplots for the returned Census data. It requires the user to specify a categorical variable and a numeric variable, and uses `PWGTP` as the weight so the plot accounts for the fact that each PUMS record may represent multiple people. 

I also made sure `year` is treated as a categorical variable when used for the x-axis, so the boxplots show separate distributions for each survey year. 

```{r}
# create a generic plot() function that 
# creates a boxplot of the numeric variable 
# at the levels of the categorical variables 

plot.census <- function(
    object,
    categorical_var,
    numeric_var
) {
  
  # make year a categorical variable 
  if (categorical_var == "year") {
    object$year <- factor(object$year)
  }
  
  ggplot(
    object,
    aes(
      x = get(categorical_var),
      y = get(numeric_var),
      weight = PWGTP
    )
  ) +
    geom_boxplot(width = 0.6) +
    labs(
      x = categorical_var,
      y = numeric_var,
      title = paste(numeric_var, "by", categorical_var)
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      axis.title = element_text(face = "bold")
    )
}
```

I tested `plot.census()` using different categorical and numeric variables. The first example compares age across sex, while the second compares age across survey years. By using different inputs, the function shows that it can be reused with different combinations of variables. 

```{r}
# test plot function

plot(
  combined_test,
  categorical_var = "SEX",
  numeric_var = "AGEP"
)

plot(
  combined_test,
  categorical_var = "year",
  numeric_var = "AGEP"
)
```

# Investigation